Meta-Analysis of the `Ironic' Effects of Intergroup Contact

### 2022-05-05 manuscript/manuscript.Rmd
update reference

Our research was motivated by the broader question whether intergroup contact helps or hinders social change. By examining the narrower question whether intergroup contact diminishes support for social change in disadvantaged groups, we tested the most prominent hypotheses explaining how intergroup contact might *hinder* social change. We did not consider whether intergroup contact might also *help* social change by reducing prejudice [for critical perspectives, see @dixon_beyond_2012; @jackman_velvet_1994] or by motivating the advantaged to acknowledge and challenge social inequality [@tropp_making_2018]. Likewise, we did not test whether, in certain conditions, positive contact with advantaged-group members might *increase* support for social change in disadvantaged groups. For example, studies found positive intergroup contact to increase collective action intentions when advantaged-group members acknowledged intergroup injustice as illegitimate [@2075] or expressed support for social change [@2033]. More broadly, intergroup contact is associated with more support for social change if it satisfies disadvantaged-group members' need for empowerment, that is, when they felt that the advantaged-group members they had contact with listened to what they had to say and perceived them as competent and intelligent [@hassler_need_2021].^[We did not consider whether, in certain conditions, positive contact might increase support for social change in disadvantaged groups because only few studies measured the relevant moderator variables and because our analyses relied on correlation matrices which are not enough to estimate interaction effects. For the latter reason, we also did not consider whether positive and negative contact might interact in their effects on support for social change [see @albzour_talking_inpress].] Further, we did not consider contact between different disadvantaged groups, which has been found to increase support for social change among the disadvantaged [@dixon_divide_2015; @93]. Therefore, this meta-analysis could not provide a complete answer to the question whether, on balance, intergroup contact helps or hinders social change.

>>>>

Our research was motivated by the broader question whether intergroup contact helps or hinders social change. By examining the narrower question whether intergroup contact diminishes support for social change in disadvantaged groups, we tested the most prominent hypotheses explaining how intergroup contact might *hinder* social change. We did not consider whether intergroup contact might also *help* social change by reducing prejudice [for critical perspectives, see @dixon_beyond_2012; @jackman_velvet_1994] or by motivating the advantaged to acknowledge and challenge social inequality [@tropp_making_2018]. Likewise, we did not test whether, in certain conditions, positive contact with advantaged-group members might *increase* support for social change in disadvantaged groups. For example, studies found positive intergroup contact to increase collective action intentions when advantaged-group members acknowledged intergroup injustice as illegitimate [@2075] or expressed support for social change [@2033]. More broadly, intergroup contact is associated with more support for social change if it satisfies disadvantaged-group members' need for empowerment, that is, when they felt that the advantaged-group members they had contact with listened to what they had to say and perceived them as competent and intelligent [@hassler_need_2021].^[We did not consider whether, in certain conditions, positive contact might increase support for social change in disadvantaged groups because only few studies measured the relevant moderator variables and because our analyses relied on correlation matrices which are not enough to estimate interaction effects. For the latter reason, we also did not consider whether positive and negative contact might interact in their effects on support for social change [see @albzour_talking_2022].] Further, we did not consider contact between different disadvantaged groups, which has been found to increase support for social change among the disadvantaged [@dixon_divide_2015; @93]. Therefore, this meta-analysis could not provide a complete answer to the question whether, on balance, intergroup contact helps or hinders social change.

### 2022-05-05 manuscript/manuscript.tex
update reference

Our research was motivated by the broader question whether intergroup
contact helps or hinders social change. By examining the narrower
question whether intergroup contact diminishes support for social change
in disadvantaged groups, we tested the most prominent hypotheses
explaining how intergroup contact might \emph{hinder} social change. We
did not consider whether intergroup contact might also \emph{help}
social change by reducing prejudice (for critical perspectives, see
Dixon et al., 2012; Jackman, 1994) or by motivating the advantaged to
acknowledge and challenge social inequality (Tropp \& Barlow, 2018).
Likewise, we did not test whether, in certain conditions, positive
contact with advantaged-group members might \emph{increase} support for
social change in disadvantaged groups. For example, studies found
positive intergroup contact to increase collective action intentions
when advantaged-group members acknowledged intergroup injustice as
illegitimate (Becker et al., 2013) or expressed support for social
change (Droogendyk et al., 2016). More broadly, intergroup contact is
associated with more support for social change if it satisfies
disadvantaged-group members' need for empowerment, that is, when they
felt that the advantaged-group members they had contact with listened to
what they had to say and perceived them as competent and intelligent
(Hässler, Ullrich, et al., 2021).\footnote{We did not consider whether,
  in certain conditions, positive contact might increase support for
  social change in disadvantaged groups because only few studies
  measured the relevant moderator variables and because our analyses
  relied on correlation matrices which are not enough to estimate
  interaction effects. For the latter reason, we also did not consider
  whether positive and negative contact might interact in their effects
  on support for social change (see Albzour et al., in press).} Further,
we did not consider contact between different disadvantaged groups,
which has been found to increase support for social change among the
disadvantaged (Dixon et al., 2015, 2017). Therefore, this meta-analysis
could not provide a complete answer to the question whether, on balance,
intergroup contact helps or hinders social change.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-albzour_talking_inpress}{}%
Albzour, M., Bady, Z., Elcheroth, E., Penic, S., Nasser, R., Reimer, N.
K., \& Green, E. G. T. (in press). Talking to a (segregation) wall:
Intergroup contact and attitudes toward normalization with {Israelis}
among {Palestinians} from the occupied territories. \emph{Political
Psychology}.

>>>>

Our research was motivated by the broader question whether intergroup
contact helps or hinders social change. By examining the narrower
question whether intergroup contact diminishes support for social change
in disadvantaged groups, we tested the most prominent hypotheses
explaining how intergroup contact might \emph{hinder} social change. We
did not consider whether intergroup contact might also \emph{help}
social change by reducing prejudice (for critical perspectives, see
Dixon et al., 2012; Jackman, 1994) or by motivating the advantaged to
acknowledge and challenge social inequality (Tropp \& Barlow, 2018).
Likewise, we did not test whether, in certain conditions, positive
contact with advantaged-group members might \emph{increase} support for
social change in disadvantaged groups. For example, studies found
positive intergroup contact to increase collective action intentions
when advantaged-group members acknowledged intergroup injustice as
illegitimate (Becker et al., 2013) or expressed support for social
change (Droogendyk et al., 2016). More broadly, intergroup contact is
associated with more support for social change if it satisfies
disadvantaged-group members' need for empowerment, that is, when they
felt that the advantaged-group members they had contact with listened to
what they had to say and perceived them as competent and intelligent
(Hässler, Ullrich, et al., 2021).\footnote{We did not consider whether,
  in certain conditions, positive contact might increase support for
  social change in disadvantaged groups because only few studies
  measured the relevant moderator variables and because our analyses
  relied on correlation matrices which are not enough to estimate
  interaction effects. For the latter reason, we also did not consider
  whether positive and negative contact might interact in their effects
  on support for social change (see Albzour et al., 2022).} Further, we
did not consider contact between different disadvantaged groups, which
has been found to increase support for social change among the
disadvantaged (Dixon et al., 2015, 2017). Therefore, this meta-analysis
could not provide a complete answer to the question whether, on balance,
intergroup contact helps or hinders social change.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-albzour_talking_2022}{}%
Albzour, M., Bady, Z., Elcheroth, E., Penic, S., Nasser, R., Reimer, N.
K., \& Green, E. G. T. (2022). Talking to a (segregation) wall:
Intergroup contact and attitudes toward normalization with {Israelis}
among {Palestinians} from the occupied territories. \emph{Political
Psychology}. \url{https://doi.org/10.1111/pops.12816}
